---
title: "Assignment 3"
subtitle: "Due at 11:59pm on October 15."
output: 
  pdf_document:
    latex_engine: xelatex 
editor: visual
---

github link: [GitHub Link](https://github.com/namo507/SURV-727)

You may work in pairs or individually for this assignment. Make sure you join a group in Canvas if you are working in pairs. Turn in this assignment as an HTML or PDF file to ELMS. Make sure to include the R Markdown or Quarto file that was used to generate it. Include the GitHub link for the repository containing these files.

So first let's load the necessary libraries
```{r}
library(xml2)
library(rvest)
library(tidyverse)
library(robotstxt)
library(knitr)
library(dplyr)
library(kableExtra)
```

## Web Scraping

In this assignment, your task is to scrape some information from Wikipedia. We start with the following page about Grand Boulevard, a Chicago Community Area.

<https://en.wikipedia.org/wiki/Grand_Boulevard,_Chicago>

The ultimate goal is to gather the table "Historical population" and convert it to a `data.frame`.

As a first step, read in the html page as an R object. Extract the tables from this object (using the `rvest` package) and save the result as a new object. Follow the instructions if there is an error. Use `str()` on this new object -- it should be a list. Try to find the position of the "Historical population" in this list since we need it in the next step.

Extract the "Historical population" table from the list and save it as another object. You can use subsetting via `[[…]]` to extract pieces from a list. Print the result.

You will see that the table needs some additional formatting. We only want rows and columns with actual values (I called the table object `pop`).

```{r}
url <- "https://en.wikipedia.org/wiki/Grand_Boulevard,_Chicago"


if(paths_allowed(url) == TRUE){
  table_url <- read_html(url)
  tables <- html_table(table_url)
}


find_census_table <- function(tables) {
  Filter(Negate(is.null), lapply(tables, function(tbl) {
    if (any(grepl("Census", 
                  gsub("\\s+", "", colnames(tbl)), 
                  ignore.case = TRUE))) {
      return(tbl)
    } else {
      return(NULL)
    }
  }))
}

census_table <- find_census_table(tables)[[1]]
census_table %>% select(,-3) %>%
  slice(-11,) %>% kable(.,type='latex')
```
Write so a brief breakdown of what this code is trying to do. It first checks if the website's paths are accessible using the paths_allowed() function. If the URL is accessible, it reads the HTML content of the page and extracts all tables using the read_html() and html_table() functions. Then, the function find_census_table() filters these tables to locate the one that includes "Census" in the column names. 

Finally, we wrote the code such that it selects the relevant table, removes the third column, excludes the eleventh row, and formats the table in LaTeX format using the kable() function for further analysis.

## Expanding to More Pages

That's it for this page. However, we may want to repeat this process for other community areas. The Wikipedia page https://en.wikipedia.org/wiki/Grand_Boulevard,\_Chicago has a section on "Places adjacent to Grand Boulevard, Chicago" at the bottom. Can you find the corresponding table in the list of tables that you created earlier? Extract this table as a new object.

Then, grab the community areas east of Grand Boulevard and save them as a character vector. Print the result.

We want to use this list to create a loop that extracts the population tables from the Wikipedia pages of these places. To make this work and build valid urls, we need to replace empty spaces in the character vector with underscores. This can be done with `gsub()`, or by hand. The resulting vector should look like this: "Oakland,\_Chicago" "Kenwood,\_Chicago" "Hyde_Park,\_Chicago"

To prepare the loop, we also want to copy our `pop` table and rename it as `pops`. In the loop, we append this table by adding columns from the other community areas.

```{r}
string_mani <- function(x){
  return(gsub(' ', '_', unlist(x, use.names = FALSE)))
}
place_adj <- tables[[4]]
place_adj <- place_adj[-c(2,4),]
east_sides <- place_adj$X3
east_sides <- string_mani(east_sides)
print(east_sides, type = 'latex')
```
This code snippet that we wrote, targets the task of obtaining community areas east of Grand Boulevard, Chicago, from a Wikipedia table. Firstly we see that it defines a function, string_mani, to replace spaces with underscores in the given input strings. It then accesses and processes the fourth table (tables[[4]]) to extract the community areas. This involves removing the second and fourth rows, followed by retrieving and transforming the entries in the third column to have underscores in place of spaces using the string_mani function. 

Finally, we print these transformed community areas as a LaTeX character vector for further use in a looping mechanism that aims to extract population tables from Wikipedia pages of these areas.

Build a small loop to test whether you can build valid urls using the vector of places and pasting each element of it after `https://en.wikipedia.org/wiki/` in a for loop. Calling `url` shows the last url of this loop, which should be `https://en.wikipedia.org/wiki/Hyde_Park,_Chicago`.

```{r}
base_url <- "https://en.wikipedia.org/wiki/"
east_sides_tables <- list(census_table)

for(i in 1:length(east_sides)){
  place_url <- paste0(base_url, east_sides[i], sep = '')
  if(paths_allowed(place_url) == TRUE){
    html_url <- read_html(place_url)
    east_tbl <- html_table(html_url)
    east_sides_tables[[i+1]] = find_census_table(east_tbl)[[1]]
  }
  else{
    print('Invalid URL')
  }
}

print(place_url)
```

```{r}
df_cleaned <- function(ltbl){
  n <- length(ltbl)
  df <- list()
  for(i in 1:n){
    df[[i]] <- ltbl[[i]][,-3]
    df[[i]] <- df[[i]][-nrow(df[[i]]),]
  } 
  return(df)
}

east_sides_tables <- df_cleaned(east_sides_tables)
print(east_sides_tables, type = 'latex')
```

Finally, extend the loop and add the code that is needed to grab the population tables from each page. Add columns to the original table `pops` using `cbind()`.

```{r}
extract_census_pop <- function(tbl) {
  tbl %>%
    select(Census, Pop.) %>%
    mutate(Pop. = as.numeric(gsub(",", "", Pop.)))
}

tables_list_clean <- lapply(east_sides_tables, 
                            extract_census_pop)

pops <- Reduce(function(x, y) full_join(x, y, by = "Census"), 
                         tables_list_clean)
pop_col_name <- c("Grand_Boulevard,_Chicago",east_sides)
colnames(pops) <- c('Census', gsub(',_Chicago','',pop_col_name))
pops <- pops %>% arrange(as.numeric(Census))
kable(pops, type='latex')
```

Ok, so in a nutshell, we coded such that, firstly it constructs URLs for each community area east of Grand Boulevard by pasting them after https://en.wikipedia.org/wiki/, testing their validity, and scraping the population tables from these pages. It then cleans the extracted tables by removing unnecessary columns and rows, and prepares them for merging. 

Next, we defined a function to extract relevant census data, then, applied this function to the cleaned tables, and combined the results into one consolidated table (pops). 

Finally, we updated column names to match the community areas and arranged the table by census data. This creates a comprehensive dataset of population figures across multiple community areas in Chicago. 

## Scraping and Analyzing Text Data

Suppose we wanted to take the actual text from the Wikipedia pages instead of just the information in the table. Our goal in this section is to extract the text from the body of the pages, then do some basic text cleaning and analysis.

First, scrape just the text without any of the information in the margins or headers. For example, for "Grand Boulevard", the text should start with, "**Grand Boulevard** on the [South Side](https://en.wikipedia.org/wiki/South_Side,_Chicago "South Side, Chicago") of [Chicago](https://en.wikipedia.org/wiki/Chicago "Chicago"), [Illinois](https://en.wikipedia.org/wiki/Illinois "Illinois"), is one of the ...". Make sure all of the text is in one block by using something like the code below (I called my object `description`).

```{r}
path_nodes <- html_elements(table_url, xpath ="//p")
path_names <- html_text(path_nodes)
print(path_names[2])
```

Using a similar loop as in the last section, grab the descriptions of the various communities areas. Make a tibble with two columns: the name of the location and the text describing the location.

Let's clean the data using `tidytext`. If you have trouble with this section, see the example shown in <https://www.tidytextmining.com/tidytext.html>

```{r}
library(tidytext)
path_names <- list()
all_path_names <- list()
for(i in 1:length(pop_col_name)){
  path_url <- paste0(base_url, pop_col_name[i], sep ='')
  path_html <- read_html(path_url)
  p_nodes <- html_elements(path_html, xpath = '//p')
  path_names[[i]] <- html_text(p_nodes)[[2]]
  all_path_names[[i]] <- html_text(p_nodes)
}

desc_df <- data.frame(name = gsub('_|,_Chicago', '', pop_col_name),
           description = do.call(rbind, path_names))
kable(desc_df, type = 'latex')
```

Create tokens using `unnest_tokens`. Make sure the data is in one-token-per-row format. Remove any stop words within the data. What are the most common words used overall?

```{r}
combined_df <- data.frame(text = unlist(all_path_names), 
                          stringsAsFactors = FALSE)
combined_tibble <- tibble(line = 1:nrow(combined_df), 
                             text = combined_df$text)
tokenized_tibble <- combined_tibble %>%
  unnest_tokens(word, text)

data("stop_words")
tidy_page <- tokenized_tibble %>%
  anti_join(stop_words)
cat("The number of words reduced: ", 
    nrow(tokenized_tibble)-nrow(tidy_page))
tidy_page %>%
  count(word, sort = TRUE)
```

Plot the most common words within each location. What are some of the similarities between the locations? What are some of the differences?

```{r}
library(ggplot2)

tidy_page %>%
  count(word, sort = TRUE) %>%
  filter(n > 10) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)
```

Right! so in this part of the code, again, we focussed on extracting and analyzing the textual content from Wikipedia pages for various Chicago community areas. Firstly, the code retrieves the main descriptive text from each page, ensuring it excludes peripheral information such as headers or sidebars. 
This text is compiled into a dataframe, with columns for the community area name and the descriptive text. Then, using the tidytext package, the text is tokenized to create a one-word-per-row format and is cleaned by removing common stop words. 

Finally, the code identifies the most frequently used words across all descriptions and visualizes these in a bar plot. This in turn offers deeper insights into the features of each community area by assisting in the comprehension of say, recurring themes and variations in the way each is described.